---
title: "hR: Toolkit for Data Analytics in Human Resources"
author: "Dale Kube"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{hR: Toolkit for Data Analytics in Human Resources}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(hR)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Transform and analyze workforce data in meaningful ways for human resources (HR) analytics. The use of three functions, `hierarchyLong`, `hierarchyWide`, and `hierarchyStats` are demonstrated below. These convert standard employee and supervisor relationship data to useful formats, summary statistics, and span of control metrics.

Install the package from CRAN by running the `install.packages("hR")` command.

## workforceHistory data

The examples in this vignette use the sample `workforceHistory` data set which reflects an artificial organization's workforce history data. The sample is reduced to a data.table containing one row per active employee and contractor in order to properly use the subsequent functions.

```{r workforceHistory}
data("workforceHistory")

# Reduce to DATE <= today to exclude future-dated records
dt = workforceHistory[DATE<=Sys.Date()]

# Reduce to max DATE and SEQ per person
dt = dt[dt[,.I[which.max(DATE)],by=.(EMPLID)]$V1]
dt = dt[dt[,.I[which.max(SEQ)],by=.(EMPLID,DATE)]$V1]

# Only consider workers who are currently active
# This provides a reliable 'headcount' data set that reflects today's active workforce
dt = dt[STATUS=="Active"]

# Exclude the CEO because she does not have a supervisor
CEO = dt[TITLE=="CEO",EMPLID]
dt = dt[EMPLID!=CEO]

# Show the prepared table
# This represents an example, active workforce
print(dt[,.(EMPLID,NAME,TITLE,SUPVID)])
```


## hierarchyLong

The `hierarchyLong` function transforms a standard set of unique employee and supervisor identifiers (employee IDs, email addresses, etc.) into an elongated format that can be used to aggregate employee data by a particular line of leadership (i.e. include everyone who rolls up to Susan). The function returns a long data.table consisting of one row per employee for every supervisor above them, up to the top of the tree. The levels represent the number of supervisors from the employee (starting with "1" for an employee's direct supervisor).

```{r hierarchyLong}
hLong = hierarchyLong(dt$EMPLID,dt$SUPVID)
print(hLong)

# Who reports up through Susan? (direct and indirect reports)
print(hLong[Supervisor==CEO])
```


## hierarchyWide

The `hierarchyWide` function transforms a standard set of unique employee and supervisor identifiers (employee IDs, email addresses, etc.) into a wide format that can be used to aggregate employee data by a particular line of leadership (i.e. include everyone who rolls up to Susan). The function returns a wide data.table with a column for every level in the hierarchy, starting from the top of the tree (i.e. "Supv1" is likely the CEO in your organization).

```{r hierarchyWide}
hWide = hierarchyWide(dt$EMPLID,dt$SUPVID)
print(hWide)

# Who reports up through Pablo? (direct and indirect reports)
print(hWide[Supv2==199827])
```

## hierarchyStats

The `hierarchyStats` function computes summary statistics and span of control metrics from a standard set of unique employee and supervisor identifiers (employee IDs, email addresses, etc.). The resulting metrics and table are accessible from a list object.

```{r hierarchyStats}
hStats = hierarchyStats(dt$EMPLID,dt$SUPVID)

# Total Levels:
print(hStats$levelsCount$value)

# Total Individual Contributors:
print(hStats$individualContributorsCount$value)

# Total People Managers:
print(hStats$peopleManagersCount$value)

# Median Direct Reports:
print(hStats$medianDirectReports$value)

# Median Span of Control (Direct and Indirect Reports):
print(hStats$medianSpanOfControl$value)

# Span of Control Table
print(hStats$spanOfControlTable)
```